Correcting for bias in high throughput sequencing data

### 2011-06-13 seqbias-paper.tex
overview figure, recreated

These biases may adversely effect transcript discovery, as low level noise may
be over reported in some regions, and in others, active transcription may be
under reported. They render untrustworthy comparisons of relative abundance
between genes, and any test of differential expression hangs on the assumption
that these biases are identical between replicates.  Additionally, in many tests
of differential expression---similarly the effects of transcript length---a higher
read count will result in higher statistical confidence. It follows that the
sensitivity of such a test will also be biased by sequence, affecting downstream
systems biology analysis such as gene ontology enrichment tests.

@@

\begin{comment}
\begin{figure*}
\begin{center}
\includegraphics[width=0.75\textwidth]{fig/freqs.pdf}
\end{center}
\caption{Nucleotide frequencies relative to the start of each read (labeled
position 0), respecting strand. The divergence of the nucleotide distribution
is summarized using the symmetrized Kullback-Leibler divergence (see Section
3).}
\label{fig:freqs}
\end{figure*}
\end{comment}

@@

\section{Method}

>>>>

These biases may adversely effect transcript discovery, as low level noise may
be over reported in some regions, and in others, active transcription may be
under reported. They render untrustworthy comparisons of relative abundance
between genes, and any test of differential expression hangs on the assuBmption
that these biases are identical between replicates.  Additionally, in many tests
of differential expression---similarly the effects of transcript length---a higher
read count will result in higher statistical confidence. It follows that the
sensitivity of such a test will also be biased by sequence, affecting downstream
systems biology analysis such as gene ontology enrichment tests.

@@

\begin{figure*}
\centerline{\includegraphics[width=0.8\textwidth]{fig/freqs.eps}}
\caption{Nucleotide frequencies are plotted relative to the start (labeled,
position 0) of each mapped read, respecting strand. The sequence is taken frmo
the genomic context surrounding the read, so that -40 to -1, for example, fall
outside the read sequence itself. The symmetrized Killback-Leibler divergence is
used to summarize the difference in nucleotide frequency compared to a fixed
estimate of background necleotide frequencies made by sampling many positions
nearby mapped reads.  Under the assumption that reads are sampled uniformly from
transcripts, each of the plots should be essentially flat.}
\label{fig:freqs}
\end{figure*}

@@

\begin{figure}
\centerline{\includegraphics[width=0.35\textwidth]{fig/overview.eps}}
\caption{An overview of the approach taken: (a) foreground sequences are sampled
from the regions surrounding the starts of mapped reads, (b) background
sequences are sampled by randomly offsetting foreground positions, (c) a
Bayesian network is classifier is trained to descriminate between the set of
sampled foreground and background sequences, (d) and the classifier is evaluated
at each position within a locus, predicting bias.}
\label{fig:overview}
\end{figure}

\section{Methods}
